The Near and Mid-infrared photometric properties of known redshift $z\geq5$ Quasars

### 2020-02-16 tex/MNRAS_submission/rv1/VHzQ_MNRAS_rv1.tex
Replaced Fig 6a, minor changes to text

WISE mapped the sky in 4 passbands, in bands centered at wavelengths
of 3.4, 4.6, 12, and 23$\mu$m. The all sky `AllWISE' catalogue
release, contains nearly 750 million detections at
high-significance\footnote{\href{wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}{wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}},
of which over 4.5M AGN candidates have been identified with 90\%
reliability \citep{Assef2018}.  \citet{Blain2013} presented WISE
mid-infrared (MIR) detections of 17 (55\%) of the then known 31
quasars at $z > 6$. However, \citet{Blain2013} was compiled with the
WISE `All-Sky' data release, as opposed to the superior AllWISE
catalogues. That sample only examined the 31 known $z>6$ quasars; our
sample has 170 objects with redshift $z \geq 6.00$ (with 108 detected
in WISE). \citet{Banados2016} reports WISE W1, W2, W3 and W4
magnitudes for the Panoramic Survey Telescope and Rapid Response
System 1 \citep[Pan-STARRS1, PS1;][]{Kaiser2002, Kaiser2010}, but with
no further investigation into the reddest WISE waveband for the
VH$z$Qs.

@@

\begin{table}
\begin{tabular}{l r r l}
\hline  \hline
Survey              & \# VH$z$Qs & (\%) & Survey reference  \\
\hline  
  ATLAS             &     4    &   ( 0.82)    &  \citet{Shanks2015} \\
  CFHQS            &   20    &   ( 4.10)    &  \citet{Willott2007} \\
  DELS$^{a}$       &   16    &   ( 3.28)    &  \citet{Dey2018} \\
  ELAIS              &     1    &   ( 0.20)    &  \citet{Vaisanen2000} \\
  FIRST              &     1    &   ( 0.20)    &  \citet{Becker1995} \\
  HSC                 &    9    &   ( 1.84)    & \citet{Miyazaki2018} \\
  IMS                 &     5    &   ( 1.02)     &  \citet{Kim2015} \\
  MMT               &   12    &   ( 2.46)     &  \citet{McGreer2013} \\
  NDWFS           &     1    &   ( 0.20)    &  \citet{JD1999} \\
  PSO                 &   100   &   (20.49)   &   \citet{Kaiser2002, Kaiser2010}
  \\
  RD                   &     1   &   ( 0.20)    &  \citet{Mahabal2005} \\
  SDSS                &  171  &    (35.04)    & \citet{EDR} \\
 SDWISE$^{b}$    &   27    &  ( 5.53)    &   \citet{WangF2016} \\
%% SHELLQs    &  $^{c}$63  &  (13.6)   &  \citet{Matsuoka2016} \\  %% if you include 8 objects from the first HSC paper...
  SHELLQs         &    55    &   (11.27)  &  \citet{Matsuoka2016}     \\  
  SUV$^{c}$       &   21     &    (4.30)  & \citet{YangJ2017} \\
  TGSS				&	1		&	(0.20)	& \citet{Saxena2018} \\
  TN				& 	1		&	(0.20)	& ?? \citet{vanBreugel1999} \\
  UHS               &    1      &  ( 0.20)     &  \citet{WangF2017} \\
  ULAS               &   10   &   ( 2.05)     & \citet{Lawrence2007} \\
  VDES$^{d}$       &   17  &    ( 3.48)     &  \citet{Reed2017} \\
  VHS                 &     1  &     ( 0.20)    & \citet{WangF2018b} \\
  VIK                 &     10    &  ( 2.05)    &  \citet{Edge2013} \\
  VIMOS           &    1      &  ( 0.20)     &   \citet{LeFevre2003} \\
  XMM				& 	1	&	 (0.20)		& \citet{Khorunzhev2017} \\
\hline  \hline
\end{tabular}
\caption{The source and number of the VH$z$Q, with the key survey reference also given. 
  Recent survey name and acronyms include: 
  $^{a}$DESI Legacy Imaging Survey; 
  $^{b}$SDWISE = SDSS+WISE; 
  $^{c}$SUV  = SDSS-ULAS/VHS; 
  $^{d}$VDES = VHS/VIKING+DES; 
%  $^{c}$Includes 8 objects with a HyperSuprimeCam \citep[HSC; ][]{Miyazaki2018} designation.
}
      \label{tab:surveys}
\end{table}

@@

We note that although we are often working close to the magnitude
limits of the surveys, Eddington bias is not a significant effect on
our measurements, since we are not discovering new sources, but
measuring properties of already known quasars. Indeed in the NIR
WFCAM/VISTA data, using forced photometry reduces the Eddington bias
to negligible levels since a flux will be measured regardless of
whether the source is just bright enough (or too faint) to be detected
above the background noise.

@@

\begin{figure*}
    \centering
%    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/DELS_J1048-0109.png} 
 %   \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
  %  \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDWISE_J0306+1853.png}
  %  \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
     \includegraphics[width=0.24\textwidth]{figures/DELS_J1048-0109.png} 
   \includegraphics[width=0.24\textwidth]{figures/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{figures/SDWISE_J0306+1853.png}
   \includegraphics[width=0.24\textwidth]{figures/SDSS_J0100+2802.png}
    \caption{VH$z$Qs with interesting MIR properties. 
      The thumbnails are the RGB colour outputs using the AllWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      {\it (a)} DELS J1048-0109 is the bluest in (W1-W2) colour quasar. 
      {\it (b)} UHS J0439+1634 was discovered by \citet{Fan2019} and is a strongly lensed quasar at $z=6.51$. 
      {\it (c)} SDWISE J0306+1853 has {\tt w3mpro} = $15.66\pm0.1$ (AB) and {\tt w4mpro } = $14.25\pm0.16$ (AB). 
      {\it (d)} SDSS J0100+2802 is the \citet{Wu2015} object.}
    \label{fig:VHzQ_W4s}
\end{figure*}
\subsection{MIR Colours}
Due to the depth all-sky coverage of the WISE (and NEOWISE-(R))
mission, several investigations have looked at how WISE detects AGN
\citep[e.g][]{Stern2012, Assef2012, Secrest2015, LaMassa2017,
Assef2018, Glikman2018, Hviding2018, LaMassa2019}

Figure~\ref{fig:W1W2W3} shows the (W1-W2) vs. (W2-W3) colour-colour
diagram for the 283 VH$z$Qs that have reported {\tt w3mpro} values
(blue to red coloured points), though we note that only 51 (18\%) of
these objects have {\tt w3snr} $>$3.0. Approximately 400,000 quasars
at redshift $z\lesssim3.0$ from the SDSS Quasar Catalog using the
Fourteenth data release \citep[DR14Q; ][]{Paris2018} are also shown
(grey colour-scale).

>>>>

WISE mapped the sky in 4 passbands, in bands centered at wavelengths
of 3.4, 4.6, 12, and 23$\mu$m. The all sky `AllWISE' catalogue
release, contains nearly 750 million detections at
high-significance\footnote{\href{wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}{wise2.ipac.caltech.edu/docs/release/allwise/expsup/sec2\_1.html}},
of which over 4.5M AGN candidates have been identified with 90\%
reliability \citep{Assef2018}.  \citet{Blain2013} presented WISE
mid-infrared (MIR) detections of 17 (55\%) of the then known 31
quasars at $z > 6$. However, \citet{Blain2013} was compiled with the
WISE `All-Sky' data release, as opposed to the superior AllWISE
catalogues. That sample only examined the 31 known $z>6$ quasars; our
sample has 174 objects with redshift $z \geq 6.00$ (with 117 detected
in WISE). \citet{Banados2016} reports WISE W1, W2, W3 and W4
magnitudes for the Panoramic Survey Telescope and Rapid Response
System 1 \citep[Pan-STARRS1, PS1;][]{Kaiser2002, Kaiser2010}, but with
no further investigation into the reddest WISE waveband for the
VH$z$Qs.

@@

\begin{table}
\begin{tabular}{l r r l}
\hline  \hline
Survey              & \# VH$z$Qs & (\%) & Survey reference  \\
\hline  
  ATLAS             &     4    &   ( 0.82)    &  \citet{Shanks2015} \\
  CFHQS            &   20    &   ( 4.10)    &  \citet{Willott2007} \\
  DELS$^{a}$       &   16    &   ( 3.28)    &  \citet{Dey2018} \\
  ELAIS              &     1    &   ( 0.20)    &  \citet{Vaisanen2000} \\
  FIRST              &     1    &   ( 0.20)    &  \citet{Becker1995} \\
  HSC                 &    9    &   ( 1.84)    & \citet{Miyazaki2018} \\
  IMS                 &     5    &   ( 1.02)     &  \citet{Kim2015} \\
  MMT               &   12    &   ( 2.46)     &  \citet{McGreer2013} \\
  NDWFS           &     1    &   ( 0.20)    &  \citet{JD1999} \\
  PSO                 &   100   &   (20.49)   &   \citet{Kaiser2002, Kaiser2010}
  \\
  RD                   &     1   &   ( 0.20)    &  \citet{Mahabal2005} \\
  SDSS                &  171  &    (35.04)    & \citet{EDR} \\
 SDWISE$^{b}$    &   27    &  ( 5.53)    &   \citet{WangF2016} \\
%% SHELLQs    &  $^{c}$63  &  (13.6)   &  \citet{Matsuoka2016} \\  %% if you include 8 objects from the first HSC paper...
  SHELLQs         &    55    &   (11.27)  &  \citet{Matsuoka2016}     \\  
  SUV$^{c}$       &   21     &    (4.30)  & \citet{YangJ2017} \\
  TGSS				&	1		&	(0.20)	& \citet{Saxena2018} \\
  TN				& 	1		&	(0.20)	& ?? \citet{vanBreugel1999} \\
  UHS               &    1      &  ( 0.20)     &  \citet{WangF2017} \\
  ULAS               &   10   &   ( 2.05)     & \citet{Lawrence2007} \\
  VDES$^{d}$       &   17  &    ( 3.48)     &  \citet{Reed2017} \\
  VHS                 &     1  &     ( 0.20)    & \citet{WangF2018b} \\
  VIK                 &     10    &  ( 2.05)    &  \citet{Edge2013} \\
  VIMOS           &    1      &  ( 0.20)     &   \citet{LeFevre2003} \\
  XMM				& 	1	&	 (0.20)		& \citet{Khorunzhev2017} \\
\hline  \hline
\end{tabular}
\caption{The source and number of the VH$z$Q, with the key survey reference also given. 
  Recent survey name and acronyms include: 
  $^{a}$DESI Legacy Imaging Survey; 
  $^{b}$SDWISE = SDSS+WISE; 
  $^{c}$SUV  = SDSS+ULAS/VHS; 
  $^{d}$VDES = VHS/VIKING+DES; 
%  $^{c}$Includes 8 objects with a HyperSuprimeCam \citep[HSC; ][]{Miyazaki2018} designation.
}
      \label{tab:surveys}
\end{table}

@@

We note that although we are often working close to the magnitude
limits of the surveys, Eddington bias does not a significant effect on
our measurements, since we are not discovering new sources, but
measuring properties of already known quasars. Indeed in the NIR
WFCAM/VISTA data, using forced photometry reduces the Eddington bias
to negligible levels since a flux will be measured regardless of
whether the source is just bright enough (or too faint) to be detected
above the background noise.

@@

\begin{figure*}
    \centering
%    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/DELS_J1048-0109.png} 
 %   \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
  %  \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDWISE_J0306+1853.png}
  %  \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
     %\includegraphics[width=0.24\textwidth]{figures/DELS_J1048-0109.png} 
     \includegraphics[width=0.24\textwidth]{figures/SDSS_J0927+2001.png}
   \includegraphics[width=0.24\textwidth]{figures/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{figures/SDWISE_J0306+1853.png}
   \includegraphics[width=0.24\textwidth]{figures/SDSS_J0100+2802.png}
    \caption{VH$z$Qs with interesting MIR properties. 
      The thumbnails are the RGB colour outputs using the AllWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      %{\it (a)} DELS J1048-0109 is the bluest in (W1-W2) colour quasar. 
      {\it (a)} SDSS J0927+2001 is the bluest in (W1-W2) colour quasar.
      {\it (b)} UHS J0439+1634 was discovered by \citet{Fan2019} and is a strongly lensed quasar at $z=6.51$. 
      {\it (c)} SDWISE J0306+1853 has {\tt w3mpro} = $15.66\pm0.1$ (AB) and {\tt w4mpro } = $14.25\pm0.16$ (AB). 
      {\it (d)} SDSS J0100+2802 is the \citet{Wu2015} object.}
    \label{fig:VHzQ_W4s}
\end{figure*}
\subsection{MIR Colours}
Due to the depth all-sky coverage of the WISE (and NEOWISE-(R))
mission, several investigations have looked at how WISE detects AGN
\citep[e.g][]{Stern2012, Assef2012, Secrest2015, LaMassa2017,
Assef2018, Glikman2018, Hviding2018, LaMassa2019}

Figure~\ref{fig:W1W2W3} shows the (W1-W2) vs. (W2-W3) colour-colour
diagram for the 330 VH$z$Qs that have reported {\tt w3mpro} values
(blue to red coloured points), though we note that only 51 (18\%) of
these objects have {\tt w3snr} $>$3.0. Approximately 400,000 quasars
at redshift $z\lesssim3.0$ from the SDSS Quasar Catalog using the
Fourteenth data release \citep[DR14Q; ][]{Paris2018} are also shown
(grey colour-scale).
